\section{Study Design and Experimental Setup}

\subsection{Controlled Simulator Setting}

The study uses two execution backends.
The \emph{primary} backend is a lightweight deterministic reference
environment implemented in pure Python; it carries the full factorial
evaluation matrix and provides the quantitative comparisons reported in the
paper. A \emph{secondary} backend runs a minimal Isaac Sim
stage~\cite{nvidiaIsaacSimDocs} on two navigation instances as a limited
qualitative cross-backend check that the same condition ordering (P0 worst,
P1/P2 succeed, R1 recovers) also appears under physics-backed execution.
The navigation family contains language-conditioned mobile-goal tasks in
indoor scenes. The manipulation family contains tabletop tasks that require
object identification, grasping, transfer, and placement. The main comparison
uses a shared easy slice plus one harder slice per family, letting us test
whether additional task burden changes the qualitative ordering or mainly
amplifies planner/tool call counts.

All conditions share the same high-level handoff:

\begin{enumerate}
\item The planner receives the current task state and the exposed tool set.
\item The planner emits an action request.
\item Under R0, the request is dispatched as issued; under R1, the runtime
validates it against the active contract and allows one repair attempt on
failure.
\item The executor runs only requests that satisfy the active dispatch interface.
\end{enumerate}

The dispatch vocabulary is family-specific and explicit. Navigation exposes
state getters and \texttt{navigate\_to}; manipulation exposes state getters
and \texttt{scripted\_pick\_place\_step}. The study therefore asks whether the
planner can hand the executor a dispatchable request under a specified
interface, not whether it can merely describe the task in natural language.

\subsection{Contract and Runtime Variants}

The controlled matrix is summarized in Table~\ref{tab:experimental-design-summary}.
P0 uses an under-specified direct-action contract, allowing the planner to
emit a task-level next action without binding it to the executor's tool
namespace. P1 uses a typed tool-call contract, requiring the planner to
choose one exposed tool and provide dispatchable arguments. P2 keeps the same
typed tool-call contract and adds a brief planner-side self-check before
emission. Table~\ref{tab:contract-interface-examples} juxtaposes one planner
emission for each contract format.

R0 is a bare-dispatch runtime: it does not perform pre-dispatch validation or
offer a repair path. R1 is a validate-and-retry runtime: it checks the
emitted request against the active contract and, on failure, returns the
validation error to the planner and allows one retry. Contract and runtime
therefore play distinct roles. Contract design can prevent invalid actions at
the source, while runtime validation can recover some contract violations that
still slip through.

\input{tables/experimental_design_summary.tex}

\subsection{Implementation Snapshot}

The study uses a repo-local deterministic planner backend implemented in
\texttt{src/isaacsim\_agent/planner/mock.py} rather than a hosted
foundation-model API. Given the current task state, exposed tools, and any
retry message, it emits one JSON action per planner call; there are therefore
no temperature, top-$p$, or rollout-sampling settings in this evaluation.
Under P0, the planner emits task-level verbs (\texttt{move\_to\_goal},
\texttt{move\_object}) that are semantically plausible but absent from the
executor's tool registry---\emph{injected boundary probes} designed to
expose the contract gap, not stochastic foundation-model errors.

Crucially, the three contract conditions separate two distinct design
axes. P0 versus P1 isolates \emph{action-interface specification}
under a fixed action-selection policy: both conditions share the same
planner decision logic, task-state input, tool registry, and planning
strategy; only the serialization format and dispatch interface change.
P2 adds one bounded planner-side self-check on top of the same typed
interface used by P1, so the P1-versus-P2 comparison isolates a
\emph{planner-side deliberation step} rather than a second interface
change.
Planner instructions for all three contracts are included in the
supplementary material and public repository so that readers can verify
this separation.

P0 realizes a direct-action JSON contract, P1 realizes typed tool JSON, and
P2 keeps the same typed tool JSON while adding a planner-side
\texttt{self\_check} string. The planner request already exposes the valid tool
names, descriptions, and required arguments through \texttt{available\_tools}.
R1 adds one more element: if parsing or validation fails, it appends the
literal error string to a short repair instruction and allows one retry. The
runtime parses the emitted JSON, validates the tool name against the
task-specific registry, enforces the exact required argument set, and then
checks task-specific payload semantics. For navigation,
\texttt{navigate\_to} must carry a valid \texttt{target\_pose} that matches the
configured goal pose. For manipulation,
\texttt{scripted\_pick\_place\_step} must carry the next scripted
\texttt{phase\_name}. The added P2 \texttt{self\_check} field is recorded in the
planner trace but does not replace runtime contract checking.

Navigation exposes \texttt{get\_robot\_state}, \texttt{get\_goal\_state}, and
\texttt{navigate\_to}. The first two are state queries; \texttt{navigate\_to}
dispatches one deterministic executor step toward the configured goal and
returns the applied target pose, progress, moved distance, and stuck-step
count. Manipulation exposes \texttt{get\_gripper\_state},
\texttt{get\_object\_state}, \texttt{get\_target\_state}, and
\texttt{scripted\_pick\_place\_step}. The first three are state queries;
\texttt{scripted\_pick\_place\_step} advances one phase of a fixed pick-and-place
script such as \texttt{move\_to\_pregrasp}, \texttt{descend\_to\_grasp}, or
\texttt{close\_gripper\_and\_attach}, and returns the applied phase plus the
updated observation and motion deltas. The executor therefore exposes
task-scoped, deterministic action primitives rather than raw controls or
free-form code execution.

\subsection{Metrics and Evaluation Protocol}

Success rate is the primary task outcome. An invalid action is an emitted
action or tool call that violates the active contract and therefore cannot be
dispatched as issued --- concretely, a non-dispatchable schema violation such
as an unknown tool name or missing required argument. This is distinct from a
semantically wrong but schema-valid action (e.g., a valid
\texttt{navigate\_to} call with an incorrect target pose), which passes
dispatch but produces an incorrect outcome, and from a task failure, which
denotes any run that does not reach the task-complete state regardless of
cause. In the structured contract conditions (P1, P2), no
semantically-wrong-but-schema-valid actions were observed in the reported
matrix; all runs in those conditions reached task-complete success.
A recovery is a run that contains an invalid action but
still ends successfully after runtime handling. Retries capture the extra
repair work used in R1. Planner calls and tool calls together define
planner/tool call counts, which is the paper's efficiency proxy.

The planner backend is deterministic by design, so the empirical unit is the
task instance under a specified condition rather than repeated stochastic
rollouts. The four main comparison cohorts cover 21 fixed task instances
(11 navigation easy, 3 manipulation easy, 4 navigation harder, and 3
manipulation harder), evaluated once under all six contract/runtime cells for
126 main-matrix runs. Two focused ablation cohorts add four task instances
each to isolate the action-interface and runtime effects, giving 29 task
instances and 146 total condition evaluations in the full fixed evaluation
set. We therefore report descriptive comparisons over this fixed evaluation
set rather than confidence intervals or significance tests across i.i.d.
stochastic replications.

The paper draws its quantitative comparisons from the main factorial results
plus targeted runtime-only, action-interface ablation, and harder-task
summaries. The combined summary provides the aggregate view, while the
slice-specific summaries support the ablations and the failure-case
interpretation. Within each slice, every compared condition is evaluated on the
same fixed tasks, which is why this task coverage still supports a controlled
systems comparison even without stochastic rollout replication.
